### 2014-08-18 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Edits to geometry, numerical mesh, and simulation parameters sections

% The analysis of the simulation results is focused on the interior conditions on the second floor before and after the interior door failure. Based on the goals of the analysis and the information collected, several engineering approximations were made. For example, the fire exhibited signs of under-ventilated conditions when firefighters arrived~\cite{NIOSH:Bowyer2}; however, the exact ignition time, fire growth rate, and heat release rate (HRR) of the fire were not known. The ignition time and growth rate were estimated to reduce the amount of required calculation time while adhering to documented observations. The HRR of the fire was estimated using documentation of the potential fuel load within the structure.

@@

The structure involved in this fire incident was a four-story residential building with a flat roof that consisted of a bitumen membrane. The exterior was primarily stucco with large windows on the top three stories on the rear side of the structure. The right side of the structure shared a common fire wall with another residential structure. The structure was built on a sloped landscape such that, from the front side of the structure, two stories were located above the street level and two stories were located below the street level. The first and second floor (i.e., the two stories above street level) had a total interior floor area of 136~m$^2$ (1,468~ft$^2$). The interior walls were composed of gypsum board over insulated wood framing. The entrance stairs located on the front side of the structure led to the front door and an interior landing, which allowed access to the first floor (10 stairs down) or the second floor (5 stairs up).

@@

The fire, which is described in Section~\ref{sec:fire}, originated in the finished living room on the basement level (near the rear side of the structure), which had a floor area of 72~m$^2$ (775~ft$^2$). More detailed discussion on the impact of the fire on the interior conditions is provided in Section~\ref{sec:model_results}. Figure~\ref{fig:post_exterior} shows the XX of the structure after the incident. Note that the majority of the damage occurred XX. Fully dimensioned drawings of the second floor, first floor, and basement are shown in Figs.~\ref{fig:drawing_second_floor}, \ref{fig:drawing_first_floor} and \ref{fig:drawing_basement}, respectively.

@@

For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}
% From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm, the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 15.5 for the 3.3~MW porch fire and 23 for the 9~MW attic fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

% As discussed in Section~\ref{sec:geometry}, the structure has dimensions of 17.15~m (56~ft) by 6.15~m (20~ft) and a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow in and out of the structure, the computational domain was extended beyond the volume of the structure. The total volume of the computational domain was 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). A grid resolution of 10~cm (3.9~in) was used, which required 2.4 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:model_parameters} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

\begin{table}[!ht]
\caption[Relevant fire model input parameters]{Relevant fire model input parameters}
\begin{tabular}{lll}
\toprule
Parameter                                 &  Description                                  &  Discussion                        \\
\midrule
Simulation Time                           &  5~min                                        &  --                                \\
Grid Cell Size                            &  10~cm                                        &  Section~\ref{sec:numerical_mesh}  \\
Ambient Temperature$^*$                   &  20~$^{\circ}$C (77~$^{\circ}$F)              &  --                                \\
Reaction: Wood~\cite{SFPE:Tewarson}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     &  Section~\ref{sec:fire}            \\
                                          &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                    \\
                                          &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                    \\
                                          &  Heat of Combustion: 16,400~kJ/kg             &                                    \\
Peak HRR                                  &  Attic Fire: 9~MW                             &  Section~\ref{sec:fire}            \\
                                          &  Porch Fire: 3.3~MW                           &                                    \\
Material: Wood~\cite{Incropera:1}         &  $k$: 0.12~\si{W/(m.K)})                      &  Section~\ref{sec:materials}       \\
                                          &  $\rho$: 510~kg/m$^3$                         &                                    \\
                                          &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                    \\
Material: Gypsum Board~\cite{WAKILI2007}  &  $k$: 0.28~\si{W/(m.K)}                       &  Section~\ref{sec:materials}       \\
                                          &  $\rho$: 810~kg/m$^3$                         &                                    \\
                                          &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                    \\
\bottomrule
\end{tabular}
\footnotesize
\\ $^{*}$ Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 7~$^{\circ}$C~\cite{NIOSH:Bowyer2}.
\normalsize
\label{tab:model_parameters}
\end{table}

@@

% Recent experiments conducted by NIST revealed similar post-fire states after doors of similar construction were exposed to fires. Figure~\ref{fig:SC_doorfold} shows doors from two different residential fire experiments in different stages of degradation. In both cases, note that the door knob assembly served as a pivot point for the upper portion of the door to collapse around. This is similar to the location of the creases in the remains of the steel door that was located between the hallway and the enclosed porch in this fire incident.

% \begin{figure}[!ht]
% \includegraphics[width=6.5in]{../Figures/SC_Door_1} \\
% \includegraphics[width=6.5in]{../Figures/SC_Door_2}
% \caption{Post-experiment images of door failures.}
% \label{fig:SC_doorfold}
% \end{figure}

>>>>

The analysis of the simulation results is focused on the interior conditions on the first floor where the downed firefighters were located and the conditions in the interior stairwell leading to the basement. Based on the goals of the analysis and the information collected, several engineering approximations were made. For example, the fire was in the early growth stage when firefighters arrived~\cite{NIOSH:Bowyer2}; however, the exact ignition time, fire growth rate, and heat release rate (HRR) of the fire were not known. The ignition time, growth rate, and HRR were estimated were estimated using documentation of the potential fuel load within the structure.

@@

The structure involved in this fire incident was a four-story residential building with a flat roof that consisted of a bitumen membrane. The exterior was primarily stucco with large windows on the top three stories on the rear side of the structure. The right side of the structure shared a common fire wall with another residential structure. The structure had a 17.73~m (58~ft) by 6.38~m (21~ft) footprint and a height of 5.9~m (19~ft), as shown in Fig.~\ref{fig:simp_geom}.

The structure was built on a sloped landscape such that, from the front side of the structure, two stories were located above the street level and two stories were located below the street level. The first and second floor (i.e., the two stories above street level) had a total interior floor area of 136~m$^2$ (1,468~ft$^2$). The interior walls were composed of gypsum board over insulated wood framing. The entrance stairs located on the front side of the structure led to the front door and an interior landing, which allowed access to the first floor (10 stairs down) or the second floor (5 stairs up).

@@

The fire, which is described in Section~\ref{sec:fire}, originated in the finished living room on the basement level (near the rear side of the structure), which had a floor area of 72~m$^2$ (775~ft$^2$). More detailed discussion on the impact of the fire on the interior conditions is provided in Section~\ref{sec:model_results}. Figure~\ref{fig:post_exterior} shows the exterior of the front and rear of the structure after the incident. Note that the majority of the exterior damage occurred above the openings located on the basement floor (where the wooden balcony is located). Fully dimensioned drawings of the second floor, first floor, and basement are shown in Figs.~\ref{fig:drawing_second_floor}, \ref{fig:drawing_first_floor} and \ref{fig:drawing_basement}, respectively.

@@

For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}
From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm, the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 18.8 for the initial couch fire (peak HRR of 5.4~MW) and 28.3 for the basement flashover fire (HRR of 15~MW). Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

As discussed in Section~\ref{sec:geometry}, the structure had exterior dimensions of 17.73~m (58~ft) by 6.38~m (21~ft) and a total height of 5.9~m (19~ft). To ensure adequately resolved fluid flow in and out of the structure, the computational domain was extended beyond the volume of the structure. The total volume of the computational domain was 17~m (56~ft) by 22~m (72~ft) by 13.6~m (45~ft). A grid resolution of 10~cm (3.9~in) was used, which required approximately 5.1 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 meshes (each containing between 289,000 and 346,800 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 3~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:model_parameters} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

\begin{table}[!ht]
\caption[Relevant fire model input parameters]{Relevant fire model input parameters}
\begin{tabular}{lll}
\toprule
Parameter                                 &  Description                                  &  Discussion                        \\
\midrule
Simulation Time                           &  9~min                                        &  --                                \\
Grid Cell Size                            &  10~cm                                        &  Section~\ref{sec:numerical_mesh}  \\
Ambient Temperature$^*$                   &  20~$^{\circ}$C (68~$^{\circ}$F)              &  --                                \\
Reaction: Wood~\cite{SFPE:Tewarson}       &  Formula: $\C\H_{1.8}\O_{0.30}\N_{0.05}$      &  Section~\ref{sec:fire}            \\
                                          &  Soot Yield: 0.131~$\mathrm{kg}/\mathrm{kg}$  &                                    \\
                                          &  CO Yield: 0.010~$\mathrm{kg}/\mathrm{kg}$    &                                    \\
                                          &  Heat of combustion: 26,200~kJ/kg             &                                    \\
Peak HRR                                  &  Initial couch fire: 5.4~MW                   &  Section~\ref{sec:fire}            \\
                                          &  Basement flashover fire: 15~MW               &                                    \\
Material: Gypsum Board~\cite{WAKILI2007}  &  $k$: 0.28~\si{W/(m.K)}                       &  Section~\ref{sec:materials}       \\
                                          &  $\rho$: 810~kg/m$^3$                         &                                    \\
                                          &  $c_{\rm p}$: 1.0~\si{kJ/(kg.K)}                  &                                    \\
\bottomrule
\end{tabular}
\footnotesize
\\ $^{*}$ Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 14~$^{\circ}$C~\cite{NIOSH:Bowyer2}.
\normalsize
\label{tab:model_parameters}
\end{table}

@@
